---
title: "NYPD_Shooting"
author: "Trent Kindvall"
date: "11/3/2021"
output:
  pdf_document: default
  html_document: default
---

###NYPD Shooting Analysis


#Libraries
```{r setup}
knitr::opts_chunk$set(echo = TRUE)
library('tidyverse')
library('ggrepel')
library('dbplyr')
```

#Data Import
```{r echo=TRUE}
df = read_csv("https://data.cityofnewyork.us/api/views/833y-fsy8/rows.csv?accessType=DOWNLOAD")
df

```

#Feature Engineering
```{r}
n <- 4  
df$YEAR <- substr(df$OCCUR_DATE, nchar(df$OCCUR_DATE) - n + 1, nchar(df$OCCUR_DATE))
```
## Frequency of shootings by Neigborhood
```{r}
boro <- table(df$BORO)
boro <- as.data.frame(boro)
boro$Percent <- round((boro$Freq / sum(boro$Freq)*100),2)

ggplot(boro, aes(x=Var1, y=Freq, fill=Var1)) + geom_bar(stat="identity") + geom_text_repel(data=boro, aes(label=Var1))

```
This graphic shows us which borough of New York had the most shootings shootings over the last 14 years. Now this graphic assumes a lot in that we are assuming that shootings are equal year in year out so now lets explore this assumption. 


To explore this further we can graph the data by year and borough.
```{r}
b_dt <- df%>%group_by(YEAR)%>%count(BORO)


dens_plot<- ggplot(data = b_dt,aes(x=YEAR, y = n ,group=BORO,colour=BORO)) + geom_line()

dens_plot

```
This graph shows that with the exception of 2020, we have been seeing a drop in almost all boroughs over the past 14 years. Staten Island is flat overall but it also has the lowest starting value so any change would be challenging. This brings us to building a model to show the correlation between shootings and year grouped by borough.
```{r}

yr<- table(df$YEAR)
yr <- as.data.frame(yr)

fit <- lm(yr$Var1~yr$Freq, data=yr)
fit
#predicted_df <- data.frame(freq_pred = predict(fit,y), Year= y)
#predicted_df

ggplot(data=b_dt, aes(x=YEAR, y=n, group=BORO,colour=BORO)) +
  geom_point()+
  geom_smooth( se = FALSE)+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```
This graph shows that while there is a general lowering of shootings over the past couple of years, in 2020 there was an increase in 2020. While this increase isn't yet a trend as the general movement of the volume of shootings is still moving downward, we could look bring in more data for 2020 as to why there was an increase to help us decide if we needed to make any changes in policing. 

```{r}

ggplot(data=yr, aes(x=Var1, y=Freq, group=1)) +
  geom_point()+
  geom_smooth(method = "lm", se = FALSE,show.legend = TRUE)+
  geom_smooth( se = FALSE,show.legend = TRUE)+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```
This graph shows us that on the whole shooting in New York City (NYC) have been dropping over the past 14 of years. Knowing that the boroughs themselves are dropping in we can definitively conclude that shootings in NYC are lowering. This model allows us to forecast forward and see 2 possible forecasting lines that we could use to predict the number of shootings in 2021 and beyond. We would need more data than we have available to decide which one we think is more likely to happen. If we looked at why the shootings rose in 2020 we might conclude that it was a single year phenomena and that the linear fit is more likely moving forward. Or we could see a change in the reason for shootings in 2020 that would would continue to increase or just tapper off at a new normal. I think that with the number of political rallys in 2020 that we could expect something in between the two in 2021 and  we start to drop again in 2022.